---
title: "Project: Climate in Guatemala"
author: "Andrea Sofía García de León"
date: "October, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is an example of an R Markdown document. It presents information about climate in Guatemala. The data was obtained from WorldClim.


## Guatemala

Guatemala is a country in Central America. Its official area is 108,889 km². The borders of Guatemala are shown on the map. 
```{r echo=FALSE, echo=FALSE, message=FALSE, warning=FALSE}

library(raster)
guatemala<-getData("GADM", country="GTM", level=2) #get country borders, other country codes can ve found in the manual
plot(guatemala)

```

Source: GADM

Guatemala is mountainous except for the south coast and the vast northern lowlands of Peten. All major cities are located in the highlands and Pacific coast regions. These regions vary in climate, elevation, and landscape, providing dramatic contrasts between humid tropical weather in the Pacific Coast and Peten, and colder, drier weather in the highland peaks.

###Precipitation

Guatemala's location between the Atlantic and Pacific Ocean makes it proned to humid tropical climate. The dinamic of rainfall is shown in the sequence.
```{r, echo=FALSE, message=FALSE, warning=FALSE}
prec<-getData("worldclim", var="prec",res=2.5) #get precipitation 
prec_ger1<- crop(prec,guatemala) #crop precipitacion to extent of guate
prec_ger2<- mask(prec_ger1, guatemala) #mask precipitacion to extent of guate
spplot(prec_ger2) #plot result

```

Source: WorldClim

As it is shown in the graph, the rainy season starts in May and ends in October. There is a canicula in July and Agoust.
```{r, echo=FALSE}
prec_avg<-cellStats(prec_ger2,stat="mean") #extract precipitation average of guate, other statistics possible as well
plot(prec_avg, 
     main="Average monthly precipitation in Guatemala", 
     type= 'b',
     xlab="Month",
     ylab="Precipitation (mm)",
     lty=4, 
     pch=16, 
     col=4)
```

###Temperature

The dinamic of mean monthly temperature is shown in the following sequence.
```{r, echo=FALSE, message=FALSE, warning=FALSE}
temp<-getData("worldclim", var="tmean",res=2.5) #get temperature
temp_ger1<- crop(temp,guatemala) #crop temp to extent of guate
temp_ger2<- mask(temp_ger1, guatemala) #mask temp to extent of guate
spplot(temp_ger2)
```

Source: WorldClim

The average mean monthly temperature is shown in the graphic. The temperature is shown in °C *10. The warm season is from March to July. The hottest month is May.

```{r,echo=FALSE }
temp_avg<-cellStats(temp_ger2,stat="mean") #extract temp average of guate, other statistics possible as well
plot(temp_avg, 
     main="Average monthly mean temperature in Guatemala", 
     type= 'b',
     xlab="Month",
     ylab="Mean Temperature (°C * 10)",
     lty=6, 
     pch=16, 
     col="orangered1")
```