---
title: "Precipitation in Guatemala"
author: "Sofía García"
date: "October, 2019"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```



## Guatemala
Guatemala is a country in Central America
Its area is 108 889 km2
```{r echo = FALSE, message=FALSE, warning=FALSE}
library(raster)
guatemala<-getData("GADM", country="GTM", level=2) #get country borders, other country codes can ve found in the manual
plot(guatemala)
```

## Precipitation
Guatemala's location between the Atlantic and Pacific Ocean makes it proned to humid tropical climate. The dinamic of rainfall is shown in the sequence.
```{r, echo=FALSE, message=FALSE, warning=FALSE}
prec<-getData("worldclim", var="prec",res=2.5) #get precipitation 
prec_ger1<- crop(prec,guatemala) #crop precipitacion to extent of guate
prec_ger2<- mask(prec_ger1, guatemala) #mask precipitacion to extent of guate
spplot(prec_ger2) #plot result
```
Source: WorldClim

##Average Precipitation Plot
As it is shown in the graph, the rainy season starts in May and ends in October. There is a canicula in July and Agoust.
```{r,echo=FALSE}
prec_avg<-cellStats(prec_ger2,stat="mean") #extract precipitation average of guate, other statistics possible as well
plot(prec_avg, 
     main="Average monthly precipitation in Guatemala", 
     type= 'b',
     xlab="Month",
     ylab="Precipitation (mm)",
     lty=4, 
     pch=16, 
     col=4)
```

##Temperature

```{r, echo=FALSE, message=FALSE, warning=FALSE}
temp<-getData("worldclim", var="tmean",res=2.5) #get temperature
temp_ger1<- crop(temp,guatemala) #crop temp to extent of guate
temp_ger2<- mask(temp_ger1, guatemala) #mask temp to extent of guate
spplot(temp_ger2)
```
Source: WorldClim

##Average mean monthly temperature
The temperature is shown in °C *10. 
```{r,echo=FALSE }
temp_avg<-cellStats(temp_ger2,stat="mean") #extract temp average of guate, other statistics possible as well
plot(temp_avg, 
     main="Average monthly mean temperature in Guatemala", 
     type= 'b',
     xlab="Month",
     ylab="Mean Temperature (°C * 10)",
     lty=6, 
     pch=16, 
     col="orangered1")
```


##Coding

```{r eval=FALSE, echo=TRUE}
install.packages("raster")
library(raster)
guatemala<-getData("GADM", country="GTM", level=2) #get country borders, other country codes can ve found in the manual
plot(guatemala)
prec<-getData("worldclim", var="prec",res=2.5) #get precipitation 
plot(prec)
prec_ger1<- crop(prec,guatemala) #crop precipitacion to extent of guate
spplot(prec_ger1) #plot results
prec_ger2<- mask(prec_ger1, guatemala) #mask precipitacion to extent of guate
spplot(prec_ger2) #plot result
prec_avg<-cellStats(prec_ger2,stat="mean") #extract precipitation average of guate, other statistics possible as well
plot(prec_avg, 
     main="Average monthly precipitation in Guatemala", 
     type= 'b',
     xlab="Month",
     ylab="Precipitation (mm)",
     lty=4, 
     pch=16, 
     col=4)
temp<-getData("worldclim", var="tmean",res=2.5) #get temperature
plot(temp)
temp_ger1<- crop(temp,guatemala) #crop temp to extent of guate
spplot(temp_ger1) #plot results
temp_ger2<- mask(temp_ger1, guatemala) #mask temp to extent of guate
spplot(temp_ger2) #plot result
temp_avg<-cellStats(temp_ger2,stat="mean") #extract temp average of guate, other statistics possible as well
plot(temp_avg, 
     main="Average monthly mean temperature in Guatemala", 
     type= 'b',
     xlab="Month",
     ylab="Mean Temperature (°C)",
     lty=6, 
     pch=16, 
     col="orangered1")
```

